---
title: "Introduction to data visualization for dental researchers"
author: Sergio Uribe
output: html_notebook
---

# Introduction

Required
```{r}
library(tidyverse)
```

```{r}
df <- read_csv("http://bit.ly/caries_visualization")
```


## The data
The structure of your data will dictate how you construct plots 

```{r}
glimpse(df)
```

```{r}
summary(df)
```

Tell R that School is a factor and not a number
```{r}
df$school <- as.character(df$school)
```

```{r}
summary(df)
```
```{r}
df %>% 
  select(-RUT)
```
```{r}
df <- df %>% 
  select(-RUT)
```

## Tables


```{r}
table(df$school, df$sex)
```

## Recode
```{r}
df %>% 
  mutate(sex = fct_recode(sex,
    "male" = "mle"
  ))
```
```{r}
df <- df %>% 
  mutate(sex = fct_recode(sex,
    "male" = "mle"
  ))
```
```{r}
table(df$school, df$sex)
```

## Reorder

```{r}
df$school <- fct_inorder(df$school)
```

```{r}
table(df$school, df$sex)
```
## Filter

```{r}
df %>% 
  filter(school == "1")
```

```{r}
df %>% 
  filter(school == "1" & sex == "female")
```
```{r}
df %>% 
  filter(D < 2)
```

```{r}
df %>% 
  filter(D > 1 & M > 1 & F > 1)
```
```{r}
df %>% 
  filter(D > 1 | M > 1 | F > 1)```


## Anscombre


## Summaries

```{r}
df %>% 
  group_by(sex) %>% 
  summarise("Age mean" = mean(age))
```


```{r}
install.packages("devtools")
library(devtools)
install_github("russHyde/exemplr")

anscombe_tidy <- exemplr::get_tidy_anscombe()
```

```{r}
anscombe_tidy
```
```{r}
anscombe_tidy %>% 
  group_by(group) %>% 
  summarise(mean_x = mean(x), sd_x = sd(x), mean_y = mean(y), sd_y = sd(y) )
```

```{r}
anscombe_tidy %>% 
  ggplot(aes(x, y, color = group)) + 
  geom_point() + 
  facet_wrap(~group) + 
  geom_smooth(method = "lm")
```

```{r}
cor(anscombe$x1, anscombe$y1)
```
```{r}
cor(anscombe$x2, anscombe$y2)
```

```{r}
anscombe_tidy %>% 
  ggplot(aes(x, y, color = group)) + 
  geom_point() + 
  facet_wrap(~group) + 
  geom_smooth(method = "loess")
```


You have two options: change by hand or omit the NA values

```{r}
df %>% 
  na.exclude(.) %>% 
  group_by(sex) %>% 
  summarise("Age mean" = mean(age, ))
```
### Ubicar al NA
```{r}
is.na(df$age)
```
```{r}
sum(is.na(df$age))
```
```{r}
which(is.na(df$age))
```

We can create a "clean of NA" dataset

```{r}
df_clean <- df[complete.cases(df),]
```

```{r}
df_clean %>% 
  group_by(sex) %>% 
  summarise("Age in months" = mean(age))
```



ggplot graphics are built step by step by adding new elements. Adding layers in this fashion allows for extensive flexibility and customization of plots.


## Aesthetics

Aesthetic mappings are the cornerstone of the grammar of graphics plotting concept. This is where the magic happens - converting continuous and categorical data into visual scales that provide access to a large amount of information in a very short time


## Plot geometry
A plot’s geometry dictates what visual elements will be used. 