---
title: "Introduction to data visualization for dental researchers"
author: Sergio Uribe
output: html_notebook
---

# Introduction

Required
```{r}
library(tidyverse)
```

```{r}
df <- read_csv("http://bit.ly/caries_visualization")
```


## The data
The structure of your data will dictate how you construct plots 

```{r}
glimpse(df)
```

```{r}
summary(df)
```

Tell R that School is a factor and not a number
```{r}
df$school <- as.character(df$school)
```

```{r}
summary(df)
```
```{r}
df %>% 
  select(-RUT)
```
```{r}
df <- df %>% 
  select(-RUT)
```

## Tables


```{r}
table(df$school, df$sex)
```

## Recode
```{r}
df %>% 
  mutate(sex = fct_recode(sex,
    "male" = "mle"
  ))
```
```{r}
df <- df %>% 
  mutate(sex = fct_recode(sex,
    "male" = "mle"
  ))
```
```{r}
table(df$school, df$sex)
```

## Reorder

```{r}
df$school <- fct_inorder(df$school)
```

```{r}
table(df$school, df$sex)
```
## Filter

```{r}
df %>% 
  filter(school == "1")
```


## Summaries

```{r}
df %>% 
  group_by(sex) %>% 
  summarise("Age mean" = mean(age))
```

You have two options: change by hand or omit the NA values

```{r}

```


ggplot graphics are built step by step by adding new elements. Adding layers in this fashion allows for extensive flexibility and customization of plots.


## Aesthetics

Aesthetic mappings are the cornerstone of the grammar of graphics plotting concept. This is where the magic happens - converting continuous and categorical data into visual scales that provide access to a large amount of information in a very short time


## Plot geometry
A plot’s geometry dictates what visual elements will be used. 